Real-Time Activity Classification by Match Filtering using Body-Worn Accelerometers

### 2016-06-17 notes/paper/paper.tex
Added more to the paper.

The matched filter is based on the cross correlation of the data $\vec{y} = \{y\}_{k=0}^{N-1}$ with the reference signal $\vec{x} = \{x\}_{k=0}^{M-1}$.
The cross correlation of $\vec{x}$ with $\vec{y}$ is defined as:
%
\begin{equation} \label{cross_correlation_eq}
(\vec{x} \star \vec{y}) = \left \{\sum_{i=0}^{M}x_{i} y_{i+k} \right \}_{k=0}^{N-M-1}
\end{equation}
%
which is simply a moving dot product.
We can then normalize \eqref{cross_correlation_eq} by the following:
%
\begin{equation} \label{norm_cross_correlation_eq}
\widehat{(\vec{x} \star \vec{y})}_k = \frac{(\vec{x} \star \vec{y})_k}{||\vec{x}|| \ || \{y_p\}_{p=k}^{M+k-1} || }, \quad k = 0,1,...,N-M-1
\end{equation}
%
where $|| \cdot ||$ is the usual $\ell_2$ norm. \eqref{norm_cross_correlation_eq} is now bounded by $[-1,1]$ and has any bias toward high energy removed.
We improve our computational efficiency by applying the convolution theorem to \eqref{cross_correlation_eq} and replacing it with:
%
\begin{equation} \label{conv_theorem}
(\vec{x} \star \vec{y}) = FFT^{-1}(FFT(\vec{x}') FFT(\vec{y})^*)
\end{equation}
%
Where $\vec{x}'$ is $\vec{x}$ that has been zero padded to the length of $\vec{y}$ with the added restriction that the length of $\vec{x}$ is no more than the length of $\vec{y}$.
We are most interested where our reference $\vec{x}$ best matches our data $\vec{y}$.
We use the resulting max correlation to define our metric for  our matched-filter output as:
%
\begin{equation} \label{matched_filter_eq}
MF_{\vec{x}}(\vec{y}) := \underset{k \in [0, M-N)}{max} \left \{\widehat{(\vec{x} \star \vec{y})}_k \right \}
\end{equation}
%
\subsection{Dimensionality Reduction}
We manage to reduce our computational resources and remove our dependence on sensor orientation at the same time through a process of dimensionality reduction.
We assume that over a short period of time, a user\textquotesingle s limb movement is constrained mostly to a two-dimensional plane.
By identifying the plane of motion, we then project the acceleration data onto this plane and re-define our $x$ and $y$ axis accordingly.
By this assumption, activities that are mostly confined to two dimensions will improve our signal identification.

If we let $\{\vec{a}_i\}_{i=0}^{N}$ represent our demeaned acceleration data over some time window $T$ in $\mathbb{R}^3$ then the covariance matrix of our acceleration data is represented by the real-symmetric $3x3$ matrix $Q = \vec{a} \vec{a}^T$.
The resulting eigen vectors of this matrix are orthogonal and point in the directions where the data varies most [cite from some textbook here].
By projecting our data on the plain defined by the two most dominant eigen vectors, we effectively reduce our representation of our data to two dimensions.
%
\subsection{Decimation}
Decimation is the first step applied to our data before any additional processing takes place.
Decimation is a process of reducing the sampling rate of a signal.
The process consists of filtering to mitigate any possible aliasing distortion effects followed by a downsampling.
There is a tradeoff between performance and throughput when we decimate our signal, therefore, we use decimation solely for the purpose of improving our throughput.

@@

If $\vec{t} = \{t_k\}_{k=0}^{N_t}$ is our training set over $N_t$ data points, we define a collection of subsets $X$ of $\vec{t}$:
%
\begin{equation} \label{X_subsets_of_training_eq}
X := \left \{ \{t_k\}_{k=p}^{p+M-1} | p=mI \right \}
\end{equation}
%
and $Y$ of $\vec{t}$:
%
\begin{equation} \label{Y_subsets_of_training_eq}
Y := \left \{ \{t_k\}_{k=p}^{p+N-1} | p=nI \right \}
\end{equation}
%
where $M \leq N$, $I$ is a positive, fixed integer, and $(n,m) \in \mathbb{Z}^2$ where $(m,n)$ vary such that $0 \leq (m,n)I$ and $(m,n)I + (M,N) - 1 \leq N_t$. Then we can determine the representative signal $\vec{s} \in X$ for our activity that satisfies the condition:
%
\begin{equation} \label{s_condition}
\sum_{y \in Y}MF_{\vec{s}}(\vec{y}) \geq \sum_{y \in Y}MF_{\vec{x}}(\vec{y}) \quad \forall \vec{x} \in X
\end{equation}

We chose this method for its simplicity and robustness against outliers in our training set. We anticipate that the user will have idle motion at the start and end of their training duration when they activate the training mode and with possible disruptions during the process.
%
% Decimation with a tent filter for throughput performance
%
\section{Performance and Throughput}
%
The Kinetisense sensor was used for the performance analysis in this paper.
This unit has a three-axis accelerometer with an acceleration range of -5g to +5g and a sampling rate of 128Hz.
Samples of walking, jogging, and bicycling have been recorded with this unit for the purpose of analyzing the performance.
A separate instance of walking data has been used specifically for training the classifier.
In figure \ref{fig:MF_performance}, \subref{fig:frac_walking_jogging} and \subref{fig:frac_walking_bicycling} show the fraction of the correlation results that exceed the given correlation threshold when comparing the walking data reference to the given activity.
As we would expect, walking data matches better to walking data than the other activities.
ROC curves were produced from these corresponding results in \subref{fig:roc_walking_jogging} and \subref{fig:roc_walking_bicycling} to illistrate the effectiveness of the matched-filter as a classifier.
Figure \ref{fig:MF_performance} shows the comparison to bicycling.
Because the bicycling motions are considerably different from walking relative to jogging, we would expect better performance from for bicycling.
%
%\begin{figure}[!tbp]
%\RawFloats
%  \centering
%     \begin{minipage}[b]{0.45\textwidth}
%        \includegraphics[angle=0,width=\textwidth]{ROC_walk_to_run.png}
%        \caption{\label{ROC_walk_to_run_fig}ROC curve with a downsample factor of 8}
%     \end{minipage}
%     \hfill
%     \begin{minipage}[b]{0.45\textwidth}
%     \includegraphics[angle=0,width=\textwidth]{ROC_walk_to_bike.png}
%        \caption{\label{ROC_walk_to_bike_fig}ROC curve with a downsample factor of 8}
%     \end{minipage}
%\caption{\label{ROC_walk_to_bike_fig}ROC curve with a downsample factor of 8}
%\end{figure}
%
\begin{figure}[!ht]
  %\captionsetup[subfigure]{labelformat=empty}
  \centering
  \subfloat[\label{fig:frac_walking_jogging}]{\includegraphics[width=.45\textwidth]{fraction_threshold_walking_to_jogging.png}}
  \quad
  \subfloat[\label{fig:frac_walking_bicycling}]{\includegraphics[width=.45\textwidth]{fraction_threshold_walking_to_bicycling.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_jogging}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_run.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_bicycling}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_bike.png}}
  \caption{Matched-filter performance comparing jogging and bicycling data against a reference for walking. Results produced with decimation 8}
  \label{fig:MF_performance}
\end{figure}
%
%\FloatBarrier
reference to figure \ref{fig:MF_performance}
%
\section{Conclusion}
conclusion text.
%
\appendices
%
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi
%
\bibliographystyle{plain}
\bibliography{reference}
\end{document}

>>>>

The matched filter is based on the cross correlation of the data $\textbf{y} = \{y\}_{k=0}^{N-1}$ with the reference signal $\textbf{x} = \{x\}_{k=0}^{M-1}$.
The cross correlation of $\textbf{x}$ with $\textbf{y}$ is defined as:
%
\begin{equation} \label{cross_correlation_eq}
(\textbf{x} \star \textbf{y}) = \left \{\sum_{i=0}^{M}x_{i} y_{i+k} \right \}_{k=0}^{N-M-1}
\end{equation}
%
which is simply a moving dot product.
We can then normalize \eqref{cross_correlation_eq} by the following:
%
\begin{equation} \label{norm_cross_correlation_eq}
\widehat{(\textbf{x} \star \textbf{y})}_k = \frac{(\textbf{x} \star \textbf{y})_k}{||\textbf{x}|| \ || \widetilde{\textbf{y}}_k || }, \quad k = 0,1,...,N-M-1
\end{equation}
%
where $|| \cdot ||$ is the usual $\ell_2$ norm and $\widetilde{\textbf{y}}_k = \{y_p\}_{p=k}^{M+k-1}$.
\eqref{norm_cross_correlation_eq} is now bounded by $[-1,1]$ and has any bias toward high energy removed.
We improve our computational efficiency by applying the convolution theorem to \eqref{cross_correlation_eq} and replacing it with:
%
\begin{equation} \label{conv_theorem}
(\textbf{x} \star \textbf{y}) = FFT^{-1}(FFT(\textbf{x}') FFT(\textbf{y})^*)
\end{equation}
%
Where $\textbf{x}'$ is $\textbf{x}$ that has been zero padded to the length of $\textbf{y}$ with the added restriction that the length of $\textbf{x}$ is no more than the length of $\textbf{y}$ and $FFT$ is the Fast Fourier Transform algorithm.
We are most interested where our reference $\textbf{x}$ best matches our data $\textbf{y}$.
We use the resulting max correlation to define our metric for  our matched-filter output as:
%
\begin{equation} \label{matched_filter_eq}
MF_{\textbf{x}}(\textbf{y}) := \underset{k \in [0, M-N)}{max} \left \{\widehat{(\textbf{x} \star \textbf{y})}_k \right \}
\end{equation}
%
If \eqref{matched_filter_eq} passes a threshold, the corresponding signal is a match.
%
\subsection{Dimensionality Reduction}
We manage to reduce our computational resources and remove our dependence on sensor orientation at the same time through a process of dimensionality reduction.
We assume that over a short period of time, a user\textquotesingle s limb movement is constrained mostly to a two-dimensional plane.
By identifying the plane of motion, we then project the acceleration data onto this plane and re-define our new axis accordingly.
By this assumption, activities that are mostly confined to two dimensions will improve our signal identification.

If we let $\textbf{a} = \{a_i\}_{i=0}^{N}$ represent our demeaned acceleration data over a specified time window in three dimensions then the covariance matrix of our acceleration data is represented by the real-symmetric $3x3$ matrix $Q = \textbf{a} \textbf{a}^T$.
The resulting eigen vectors of this matrix are orthogonal and point in the directions where the data varies most [cite from some textbook here].
By projecting our data on the plain defined by the two most dominant eigen vectors, we effectively reduce our representation of our data to two dimensions.

To compute the matched-filter output for two dimensions, we generalize \eqref{matched_filter_eq} by the following:
%
\begin{equation} \label{cross_correlation_eq_2}
\widehat{(\textbf{x}^{1,2} \star \textbf{y}^{1,2})}_k^2 = \frac{(\textbf{x}^1 \star \textbf{y}^1)_k + (\textbf{x}^2 \star \textbf{y}^2)_k}{||[ \textbf{x}^1, \textbf{x}^2 ]|| \ || [ \widetilde{\textbf{y}}_k^1, \widetilde{\textbf{y}}_k^2 ] || },
\end{equation}
%
\begin{equation} \label{matched_filter_eq_2}
MF_{\textbf{x}}^2(\textbf{y}) := \underset{k \in [0, M-N)}{max} \left \{\widehat{(\textbf{x} \star \textbf{y})}^2_k \right \}
\end{equation}
%
for $ k = 0,1,...,N-M-1 $.\\
%
This works because both dimensions are always in phase with each other for a given activity
\subsection{Decimation}
Decimation is the first step applied to our data before any additional processing takes place.
Decimation is a process of reducing the sampling rate of a signal.
The process consists of filtering to mitigate any possible aliasing distortion effects followed by a downsampling.
There is a tradeoff between performance and throughput when we decimate our signal, therefore, we use decimation solely for the purpose of improving our throughput.

@@

If $\textbf{t} = \{t_k\}_{k=0}^{N_t}$ is our training set over $N_t$ data points, we define a collection of subsets $X$ of $\textbf{t}$:
%
\begin{equation} \label{X_subsets_of_training_eq}
X := \left \{ \{t_k\}_{k=p}^{p+M-1} | p=mI \right \}
\end{equation}
%
and $Y$ of $\textbf{t}$:
%
\begin{equation} \label{Y_subsets_of_training_eq}
Y := \left \{ \{t_k\}_{k=p}^{p+N-1} | p=nI \right \}
\end{equation}
%
where $M \leq N$, $I$ is a positive, fixed integer, and $(n,m) \in \mathbb{Z}^2$ where $(m,n)$ vary such that $0 \leq (m,n)I$ and $(m,n)I + (M,N) - 1 \leq N_t$. Then we can determine the representative signal $\textbf{s} \in X$ for our activity that satisfies the condition:
%
\begin{equation} \label{s_condition}
\sum_{y \in Y}MF_{\textbf{s}}(\textbf{y}) \geq \sum_{y \in Y}MF_{\textbf{x}}(\textbf{y}) \quad \forall \textbf{x} \in X
\end{equation}

We chose this method for its simplicity and robustness against outliers in our training set. We anticipate that the user will have idle motion at the start and end of their training duration when they activate the training mode and with possible disruptions during the process.
%
% Decimation with a tent filter for throughput performance
%
\section{Performance and Throughput}
%
The Kinetisense sensor was used for the performance analysis in this paper.
This unit has a three-axis accelerometer with an acceleration range of -5g to +5g and a sampling rate of 128Hz.
Samples of walking, jogging, and bicycling have been recorded with this unit for the purpose of analyzing the performance.
A separate instance of walking data has been used specifically for training the classifier.
In figure \ref{fig:MF_performance}, \subref{fig:frac_walking_jogging} and \subref{fig:frac_walking_bicycling} show the fraction of the correlation results that exceed the given correlation threshold when comparing the walking data reference to the given activity.
As we would expect, walking data matches better to walking data than the other activities.
ROC curves were produced from these corresponding results in \subref{fig:roc_walking_jogging} and \subref{fig:roc_walking_bicycling} to illistrate the effectiveness of the matched-filter as a classifier.
Figure \ref{fig:MF_performance} shows the comparison to bicycling.
Because the bicycling motions are considerably different from walking relative to jogging, we would expect better performance from bicycling.
%
%\begin{figure}[!tbp]
%\RawFloats
%  \centering
%     \begin{minipage}[b]{0.45\textwidth}
%        \includegraphics[angle=0,width=\textwidth]{ROC_walk_to_run.png}
%        \caption{\label{ROC_walk_to_run_fig}ROC curve with a downsample factor of 8}
%     \end{minipage}
%     \hfill
%     \begin{minipage}[b]{0.45\textwidth}
%     \includegraphics[angle=0,width=\textwidth]{ROC_walk_to_bike.png}
%        \caption{\label{ROC_walk_to_bike_fig}ROC curve with a downsample factor of 8}
%     \end{minipage}
%\caption{\label{ROC_walk_to_bike_fig}ROC curve with a downsample factor of 8}
%\end{figure}
%
\begin{figure}[!ht]
  %\captionsetup[subfigure]{labelformat=empty}
  \centering
  \subfloat[\label{fig:frac_walking_jogging}]{\includegraphics[width=.45\textwidth]{fraction_threshold_walking_to_jogging.png}}
  \quad
  \subfloat[\label{fig:frac_walking_bicycling}]{\includegraphics[width=.45\textwidth]{fraction_threshold_walking_to_bicycling.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_jogging}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_run.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_bicycling}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_bike.png}}
  \caption{Matched-filter performance comparing jogging and bicycling data against a reference for walking. Results produced with decimation 8}
  \label{fig:MF_performance}
\end{figure}
%
%\FloatBarrier
reference to figure \ref{fig:MF_performance}
%
\section{Conclusion}
conclusion text.
%
\appendices
%
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi
%
\bibliographystyle{plain}
\bibliography{reference}
\end{document}
